---
title: "Basic Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(poissonsuperlearner)
library(data.table)
library(riskRegression)
```

# Introduction

This vignette gives fast examples of the Poisson Super Learner workflow after
the refactoring. It focuses on:

* fitting `Superlearner()` with one learner library shared by all causes;
* fitting `Superlearner()` with one learner library per cause;
* using `summary()` for the fitted super learner;
* using `predictRisk()` for the same model selectors.

The examples use small simulated data, two folds, and simple `glmnet` learners
with fixed `lambda` values so the vignette remains quick to run during package
checks.

# Data

We simulate a small competing-risks data set. The observed follow-up time is
stored in `time` and the event indicator in `event`, where `0` denotes
censoring, `1` denotes cardiovascular disease, and `2` denotes death without
prior cardiovascular disease.

```{r}
d <- simulateStenoT1(
  n = 45,
  scenario = "alpha",
  competing_risks = TRUE,
  seed = 1
)

d <- d[, .(
  id,
  time,
  event,
  sex,
  age,
  diabetes_duration,
  value_LDL,
  value_Smoking
)]

head(d)
```

# One Learner Library For All Causes

A learner library is a list of initialized learner objects. If a single library
is supplied in a competing-risks analysis, the same library is used for all
causes.

```{r}
shared_library <- list(
  simple = Learner_glmnet(
    covariates = c("sex", "diabetes_duration"),
    cross_validation = FALSE,
    lambda = 0
  ),
  shrink = Learner_glmnet(
    covariates = c("sex", "age", "value_LDL"),
    cross_validation = FALSE,
    lambda = 0.05,
    alpha = 1
  )
)

fit_shared <- Superlearner(
  data = d,
  id = "id",
  status = "event",
  event_time = "time",
  learners = shared_library,
  number_of_nodes = 3,
  nfold = 2
)
```

`summary()` gives a compact overview of the fitted super learner, including the
number of causes, retained learner labels, cross-validated deviances, and
meta-learner coefficients when a meta-learner was fitted.

```{r}
summary(fit_shared)
```

`predictRisk()` returns one row per subject and one column per requested
prediction time. The `model` argument uses the same selectors: `"sl"` for the
stacked ensemble, `"discrete_sl"` for the best cross-validated base learner per
cause, or learner labels such as `"simple"` and `"shrink"` for models stored in
the ensemble.

```{r}
newdata <- d[1:2]
times <- c(1, 2)

risk_shared_sl <- predictRisk(
  fit_shared, newdata = newdata, times = times, cause = 1, model = "sl"
)

risk_shared_discrete <- predictRisk(
  fit_shared, newdata = newdata, times = times, cause = 1, model = "discrete_sl"
)

risk_shared_simple <- predictRisk(
  fit_shared, newdata = newdata, times = times, cause = 1, model = "simple"
)

risk_shared_shrink <- predictRisk(
  fit_shared, newdata = newdata, times = times, cause = 1, model = "shrink"
)

list(
  sl = risk_shared_sl,
  discrete_sl = risk_shared_discrete,
  simple = risk_shared_simple,
  shrink = risk_shared_shrink
)
```

# One Learner Library Per Cause

For competing risks, `learners` can also be a list with one learner library per
cause. This allows different covariates, tuning parameters, or labels for each
cause.

```{r}
libraries_by_cause <- list(
  cvd = list(
    cvd_simple = Learner_glmnet(
      covariates = c("sex", "diabetes_duration"),
      cross_validation = FALSE,
      lambda = 0
    ),
    cvd_shrink = Learner_glmnet(
      covariates = c("age", "value_LDL"),
      cross_validation = FALSE,
      lambda = 0.05,
      alpha = 1
    )
  ),
  death = list(
    death_simple = Learner_glmnet(
      covariates = c("sex", "age"),
      cross_validation = FALSE,
      lambda = 0
    ),
    death_shrink = Learner_glmnet(
      covariates = c("diabetes_duration", "value_Smoking"),
      cross_validation = FALSE,
      lambda = 0.05,
      alpha = 1
    )
  )
)

fit_by_cause <- Superlearner(
  data = d,
  id = "id",
  status = "event",
  event_time = "time",
  learners = libraries_by_cause,
  number_of_nodes = 3,
  nfold = 2
)
```

The fitted object can be summarized in the same way.

```{r}
summary(fit_by_cause)
```

The stacked and discrete super learner selectors still work as scalar model
selectors for prediction.

```{r}
risk_by_cause_sl <- predictRisk(
  fit_by_cause, newdata = newdata, times = times, cause = 1, model = "sl"
)

risk_by_cause_discrete <- predictRisk(
  fit_by_cause, newdata = newdata, times = times, cause = 1, model = "discrete_sl"
)

list(
  sl = risk_by_cause_sl,
  discrete_sl = risk_by_cause_discrete
)
```

When selecting learners from cause-specific libraries, provide one selector per
cause. The first entry selects the learner for cause 1 and the second entry
selects the learner for cause 2.

```{r}
cause_specific_model <- c("cvd_simple", "death_shrink")
cause_specific_model_alt <- c("cvd_shrink", "death_simple")

risk_by_cause_selected <- predictRisk(
  fit_by_cause, newdata = newdata, times = times, cause = 1,
  model = cause_specific_model
)

risk_by_cause_selected_alt <- predictRisk(
  fit_by_cause, newdata = newdata, times = times, cause = 1,
  model = cause_specific_model_alt
)

list(
  selected_learners = risk_by_cause_selected,
  selected_learners_alt = risk_by_cause_selected_alt
)
```

Integer selectors are also supported. For cause-specific libraries, an integer
vector can choose different retained learner positions for different causes.

```{r}
risk_by_cause_indexed <- predictRisk(
  fit_by_cause, newdata = newdata, times = times, cause = 1, model = c(1, 2)
)

risk_by_cause_indexed
```
