---
title: "Basic Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(poissonsuperlearner)
library(riskRegression)
```

# Introduction

This vignette gives a minimal example of how to use `poissonsuperlearner` to fit
piecewise-constant hazard models and combine them in a Poisson super learner.
The package works by expanding subject-level time-to-event data to a long
Poisson format on a grid of time intervals, fitting one or more base learners to
the corresponding cause-specific hazards, and optionally combining them through a
meta-learner.

The same workflow can be used both for a single fitted learner via
`fit_learner()` and for an ensemble via `Superlearner()`.

# Generate synthetic Steno data

We first simulate a small synthetic data set with `simulateStenoT1()`. The
simulator is inspired by the Steno Type-1 risk engine and generates correlated
baseline covariates such as age, diabetes duration, LDL, systolic blood
pressure, HbA1c, albuminuria, eGFR, smoking, and physical activity. Event times
are then generated from Weibull proportional hazards models. In scenario
`"alpha"`, the event mechanism is linear in the covariates, and with
`competing_risks = TRUE` the data contain two causes: cardiovascular disease
(CVD) and death without prior CVD. In scenario `"beta"`, the CVD hazard instead
includes additional nonlinear hinge-squared effects of age and LDL.

Here we use the `"alpha"` scenario with competing risks to illustrate the
cause-specific output of the package on a subset of the covariates.

```{r}
set.seed(42)
d <- simulateStenoT1(
  n = 150,
  scenario = "alpha",
  competing_risks = TRUE
)

head(d[, .(id, time, event, age, diabetes_duration, value_LDL, sex)])
```

The observed follow-up time is stored in `time` and the event indicator in
`event`. With competing risks, `event = 0` denotes censoring, `event = 1`
corresponds to CVD, and `event = 2` to death without prior CVD.

# Define a small learner library

Each learner models the piecewise-constant hazard on a grid of time intervals.
In the examples below, the grid is controlled by `number_of_nodes = 5`, which
creates a quantile-based time grid used internally for the long Poisson
representation.

We define two simple learners based on `glmnet`: one unpenalized Poisson model
and one lasso-type learner.

```{r}
lglmnet0 <- Learner_glmnet(
  covariates = c("sex", "diabetes_duration"),
  cross_validation = FALSE,
  lambda = 0
)

lglmnet1 <- Learner_glmnet(
  covariates = c("value_Smoking", "value_LDL"),
  cross_validation = TRUE,
  alpha = 1
)

learners_list <- list(
  glm = lglmnet0,
  lasso = lglmnet1
)
```

# Alternative learner definitions and wrapper arguments

The package currently provides three learner classes: `Learner_glmnet()`,
`Learner_gam()`, and `Learner_hal()`. The code below is shown for
illustration only and is therefore not run. It highlights how learners are
defined by initializing Reference Class objects and then passing them to
`fit_learner()` or `Superlearner()`.

```{r, eval = FALSE}
l_glmnet <- Learner_glmnet(
  covariates = c("sex", "diabetes_duration", "value_LDL"),
  cross_validation = TRUE,
  alpha = 1
)

l_gam <- Learner_gam(
  covariates = c("s(age)", "s(value_LDL)", "sex"),
  method = "fREML",
  discrete = TRUE
)

l_hal <- Learner_hal(
  covariates = c("age", "diabetes_duration", "value_LDL"),
  cross_validation = TRUE,
  num_knots = c(10L, 5L),
  max_degree = 2L
)
```

Some learner arguments are specific to the piecewise-constant hazard workflow
implemented in `poissonsuperlearner`. In particular, `covariates` defines the
terms used in the long-format hazard model. The current learners add the
interval-specific node effects that represent the baseline hazard internally
from the chosen time grid, so there is no separate `add_nodes` argument to set.
For the HAL learner, `num_knots`, `max_degree`, and `maxit_prefit` also control
the package-specific basis construction and screening step used before the
final penalized fit.

A useful feature of the package is that, except for HAL, the learners mainly
wrap existing fitting routines. `Learner_glmnet()` forwards additional
arguments in `...` to `glmnet::glmnet()` or `glmnet::cv.glmnet()`, and
`Learner_gam()` forwards them to `mgcv::bam()`. This means that most tuning can
be done with the familiar arguments from the underlying packages rather than
through a separate package-specific interface.

For GAMs, smooth terms are specified directly in the `covariates` vector using
standard `mgcv` syntax. For example, if a spline effect of age is desired, one
simply writes `"s(age)"`; tensor products or other `mgcv` terms can be passed in
the same way.

The HAL learner is more specialized. It includes several parameters that are
specific to the piecewise-constant hazard implementation, such as the number of
basis cutpoints per interaction order (`num_knots`) and the maximum interaction
degree (`max_degree`), while still relying on the underlying `glmnet` fitting
routine for the penalized Poisson regression step. In other words, HAL combines
a package-specific basis expansion with the familiar `glmnet` optimization
controls available through `...`.

# Fit an ensemble and individual learners

`Superlearner()` fits all learners, obtains cross-validated predictions for the
meta-learning step, and then fits one cause-specific meta-learner per cause.
For comparison, `fit_learner()` fits a single learner directly.

`Superlearner()` and `fit_learner()` can be supplied either with an explicit
grid of time nodes `(nodes)` or with the number of nodes selected
from the observed quantiles (`number_of_nodes`) used to construct that grid;
`nfold` specifies the number of folds used for internal cross-validation;
if an `id` variable is provided it identifies
which rows belong to the same individual, whereas
if `id` is omitted each row is treated as a separate observation;
the `status` variable must be coded as `0` for censoring and positive integers
for the competing event types; and `event_time` names the event or
follow-up time variable on which the model is built.



```{r}
sl_model <- Superlearner(
  d,
  id = "id",
  status = "event",
  event_time = "time",
  learners = learners_list,
  number_of_nodes = 5,
  nfold = 3
)

```

The same learner library is used for both causes above. For competing risks,
`learners` can alternatively be a list with one learner library per cause. Here
cause 1 (CVD) and cause 2 (death without prior CVD) get different learner
specifications. The example is not evaluated to keep the vignette fast.

```{r, eval = FALSE}
cvd_library <- list(
  cvd_clinical = Learner_glmnet(
    covariates = c("age", "diabetes_duration", "value_LDL"),
    cross_validation = FALSE,
    lambda = 0
  ),
  cvd_lasso = Learner_glmnet(
    covariates = c("age", "value_LDL", "value_HbA1c"),
    cross_validation = TRUE,
    alpha = 1
  )
)

death_library <- list(
  death_clinical = Learner_glmnet(
    covariates = c("age", "sex", "value_eGFR"),
    cross_validation = FALSE,
    lambda = 0
  ),
  death_gam = Learner_gam(
    covariates = c("s(age)", "sex", "s(value_eGFR)"),
    method = "fREML",
    discrete = TRUE
  )
)

cause_specific_libraries <- list(
  cvd = cvd_library,
  death_without_cvd = death_library
)

sl_model_by_cause <- Superlearner(
  d,
  id = "id",
  status = "event",
  event_time = "time",
  learners = cause_specific_libraries,
  number_of_nodes = 5,
  nfold = 3
)
```

```{r}
l0_model <- fit_learner(
  d,
  id = "id",
  learner = lglmnet0,
  status = "event",
  event_time = "time",
  number_of_nodes = 5
)

l1_model <- fit_learner(
  d,
  id = "id",
  learner = lglmnet1,
  status = "event",
  event_time = "time",
  number_of_nodes = 5
)
```

# Inspect the fitted objects

The fitted objects have `print()`, `summary()`, and `coef()` methods.

`print()` is useful for a quick look at the stored fitted object. For a fitted
super learner, you can print the stacked meta-learner or one of the stored base
learners.

```{r}
print(sl_model, cause = 1)
print(sl_model, cause = 1, model = "glm")
print(l0_model, cause = 1)
```

`summary()` provides a more informative overview. For a fitted super learner, it
prints a compact description of the ensemble, the number of causes, learners,
folds and nodes, the cross-validated Poisson deviance used to compare the base
learners, and the cause-specific meta-learner coefficients. These coefficients
show how the ensemble combines the learners for each cause.

```{r}
summary(sl_model)
```

You can also summarize one of the base learners stored inside the ensemble, or a
learner fitted directly with `fit_learner()`.

```{r}
summary(sl_model, cause = 1, model = "glm")
summary(l0_model, cause = 1)
```

`coef()` extracts model coefficients. For a fitted super learner, `coef()`
returns the meta-learner coefficients by default, while the `model` argument can
be used to extract coefficients from one of the stored base learners.

```{r}
coef(sl_model, cause = 1)
coef(sl_model, cause = 1, model = "glm")
coef(l0_model, cause = 1)
```

# Predict hazards, survival, and absolute risk

The low-level `predict()` method returns a data set with one row per requested
prediction time. In addition to the original covariates, it includes the
predicted cause-specific piecewise hazards (`pwch_1`, `pwch_2`, ...), the
predicted survival function, and the absolute risk for the chosen cause.

```{r}
pred_sl <- predict(
  sl_model,
  newdata = d[1:2],
  times = c(2, 5),
  cause = 1
)

pred_sl
```

For interoperability with `riskRegression`, the package also provides
`predictRisk()`. This returns a matrix with one column per requested time.

Risk predictions can be obtained directly from a learner fitted with
`fit_learner()`:

```{r}
cbind(
  glm_direct = predictRisk(l0_model, newdata = d[1, ], times = 5, cause = 1),
  lasso_direct = predictRisk(l1_model, newdata = d[1, ], times = 5, cause = 1)
)
```

The same is possible from the fitted super learner object. Setting
`model = "sl"` uses the ensemble prediction, while setting `model` equal to a
learner label such as `"glm"` or `"lasso"` uses the corresponding stored base
learner from the fitted ensemble.

```{r}
cbind(
  sl = predictRisk(sl_model, newdata = d[1, ], times = 5, cause = 1, model = "sl"),
  glm_from_ensemble = predictRisk(sl_model, newdata = d[1, ], times = 5, cause = 1, model = "glm"),
  lasso_from_ensemble = predictRisk(sl_model, newdata = d[1, ], times = 5, cause = 1, model = "lasso")
)
```

This illustrates the two main prediction workflows supported by the package:
using a single fitted learner with `fit_learner()` or using the fitted ensemble
returned by `Superlearner()`.


