---
title: "Manuscript Replication Material"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Manuscript Replication Material}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
knitr::opts_chunk$set(eval = FALSE, include = FALSE)
library(tmlensemble)
library(survival)
library(prodlim)
library(riskRegression)
library(pammtools)
```

```{r eval=FALSE, include=FALSE}
# load the simulator 

synthesize_TD1_riskengine_data <- function(n,keep = NULL){
    #
    # empty model
    #
    sim_model <- lava::lvm()
    #
    # covariates
    #
    lava::distribution(sim_model,~sex) <- lava::binomial.lvm(p=0.462559634930512)
    lava::distribution(sim_model,~age) <- lava::normal.lvm(mean=42.2853710165578,sd=16.1358967048911)
    # this parameter seems to affect the age range:
    lava::covariance(sim_model,~age) <- 40
    lava::distribution(sim_model,~diabetes_duration) <- lava::normal.lvm(mean=-0.721917658912036,sd=11.8217218392295)
    lava::distribution(sim_model,~value_SBP) <- lava::normal.lvm(mean=113.015053986869,sd=15.1994730394504)
    lava::distribution(sim_model,~value_LDL) <- lava::normal.lvm(mean=2.34380835266364,sd=0.78907040283751)
    lava::distribution(sim_model,~value_HBA1C) <- lava::normal.lvm(mean=66.1642378865059,sd=15.1450866080122)
    lava::distribution(sim_model,~log2_eGFR_young) <- lava::normal.lvm(mean=-10.7725954841897,sd=1.85777815372659)
    lava::distribution(sim_model,~log2_eGFR_old) <- lava::normal.lvm(mean=-7.16808301118136,sd=0.295067430820889)
    lava::distribution(sim_model,~value_Smoking) <- lava::binomial.lvm(p=0.00406502756222992)
    lava::distribution(sim_model,~value_Motion) <- lava::binomial.lvm(p=0.647591503001192)
    # albuminuria
    sim_model <- lava::categorical(sim_model,~value_Albuminuria,labels=c('Normal','Micro','Macro'),K=3,p=c(0.839037544077992,0.121344119477287))
    lava::transform(sim_model, value_AlbuminuriaMicro~value_Albuminuria) <- function(x){1*(c(x)=='Micro')}
    lava::transform(sim_model, value_AlbuminuriaMacro~value_Albuminuria) <- function(x){1*(c(x)=='Macro')}
    #
    # outcome
    #
    # latent event times: .1 is cvd .2 death without cvd .0 is censoring
    lava::distribution(sim_model,~time.event.1) <- lava::coxWeibull.lvm(scale=0.00344127508109558,shape=1.1383691898171)
    lava::distribution(sim_model,~time.event.2) <- lava::coxWeibull.lvm(scale=8.8338885667654e-05,shape=1.49710154034773)
    lava::distribution(sim_model,~time.event.0) <- lava::coxWeibull.lvm(scale=0.0105086718102263,shape=2.2578792372401)
    # observed event time
    sim_model <- lava::eventTime(sim_model,time_cvd ~ min(time.event.0=0, time.event.1=1, time.event.2=2),'status_cvd')
    # uncensored event time
    sim_model <- lava::eventTime(sim_model,uncensored_time_cvd ~ min(time.event.1=1, time.event.2=2),'uncensored_status_cvd')
    #
    # structural equations
    #
    ## lava::regression(sim_model,age~sex) <- c(-0.605030442652732)
    lava::regression(sim_model,age~sex) <- c(-0.105030442652732)
    lava::regression(sim_model,diabetes_duration~age+sex) <- c(0.461453096059993,1.8834280372868)
    lava::regression(sim_model,value_SBP~diabetes_duration+age+sex) <- c(0.173711840258398,0.408090766933998,-5.26413576607228)
    lava::regression(sim_model,value_LDL~value_SBP+diabetes_duration+age+sex) <- c(0.0017997232707461,-0.00215485947423231,-0.00194439878409113,-0.0247941295129788)
    lava::regression(sim_model,value_HBA1C~value_LDL+value_SBP+diabetes_duration+age+sex) <- c(2.41425803932559,-0.0366520181671025,0.105793793763282,-0.113517351234731,1.14432850532257)
    lava::regression(sim_model,log2_eGFR_young~value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.0378605757282762,0.369860071363398,-0.000381556996934381,0.0759866855173213,-0.000921667597434942,0.0109443806788427,0.173985375155216,-0.05148192662188)
    # Fixme: it is odd to let the log2_eGFR_old depend on log2_eGFR_young
    lava::regression(sim_model,log2_eGFR_old~log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.998329448050345,0.13212871605835,0.627054487201391,-0.0017321066682523,-0.0168673394900242,0.000656653716889479,0.00231859330125351,0.0105241259989393,0.0666215469676934)
    lava::regression(sim_model,value_Smoking~log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.573224437209381,-0.533358383827049,0.417778953216348,0.811962094950751,0.0258095967578417,0.128112260674147,-0.0102291702067243,-0.00575108475194562,0.00529875681026287,-0.130756787272673)
    lava::regression(sim_model,value_Motion~value_Smoking+log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.31142087785739,-0.134144429182559,-0.0862952685994639,-0.121332306983678,-0.334529806176446,-0.0100110418345244,-0.138145721076664,0.00690508520692886,-0.0080292376964733,-0.00517733988644732,0.070748099701443)
    lava::regression(sim_model,time.event.0~value_Motion+value_Smoking+log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.0389267779620385,-0.106325304898881,-0.137101392329058,-0.118521604228187,-0.0330093143101979,0.0763711547050245,-0.00617576911315076,0.0509424048213361,-0.00757123095797519,-0.0125850223581346,-0.0130146695831196,-0.029633759327063)
    lava::regression(sim_model,time.event.1~value_Motion+value_Smoking+log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.176043482185139,0.311354666798778,0.442184803104845,0.480004039250349,0.458766977417934,0.760940706393913,0.0143694959616522,0.0701435790868005,0.00607734573133048,0.00980515944727106,0.0412496430494417,-0.224377188536924)
    lava::regression(sim_model,time.event.2~value_Motion+value_Smoking+log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.410553409305998,0.833815403131808,0.3289175620287,0.313118375461698,0.225523885897843,0.604004891420425,0.0102472549995989,0.0287051547482761,0.00156423178426314,0.00678536173681053,0.0803874049498844,-0.276186036048047)
    # construction of eGFR

    # browser()
    lava::transform(sim_model,eGFR~log2_eGFR_old+log2_eGFR_young+age) <- function(x){as.numeric(x[["age"]]<40)*100*2^{x[["log2_eGFR_young"]]}+as.numeric(x[["age"]]>40)*100*2^{x[["log2_eGFR_old"]]}}
    # simulate from model
    d <- lava::sim(sim_model,n)
    data.table::setDT(d)
    d[,log2_eGFR_old := NULL]
    d[,log2_eGFR_young := NULL]
    d[,value_AlbuminuriaMicro := NULL]
    d[,value_AlbuminuriaMacro := NULL]
    d[,time.event.0 := NULL]
    d[,time.event.1 := NULL]
    d[,time.event.2 := NULL]
    if (length(keep)>0) d[,keep,with=FALSE]
    else d[]
}



```


# Replicating Cox 

## Example on pammtools real data (survival problem)

Import data and remove ties.

```{r eval=FALSE, include=FALSE}

veteran <- survival::veteran
# remove ties
vetu <- filter(veteran, !duplicated(time))
setDT(vetu)
vetu[,id:=1:nrow(vetu)]
## split data in test and training
vetu_train <- vetu[1:70,]
vetu_test <- vetu[71:100,]

```

```{r eval=FALSE, include=FALSE}
cph_age <- coxph(Surv(time, status) ~ age, data = vetu_train,x=TRUE)
coef(cph_age)

## gam model

lgam <- Learner_gam(covariates = c("age"),
                    intercept=FALSE)


outgam <- Superlearner(vetu_train,
                     id="id",
                     status="status",
                     event_time = "time",
                     learners=list(lgam
                     ))
```

Estimates are correct.

```{r eval=FALSE, include=FALSE}
coef(cph_age)

## US
head(coef(outgam$superlearner[[1]]$learners_fit))[['age']]
```

As expected, same risk predictions for observed times.

```{r eval=FALSE, include=FALSE}

{
  set.seed(42)
  some_times <- sample(vetu_train$time,3)
}

a<-predictRisk(cph_age,newdata=vetu_test,
               times=some_times,cause =1)

b<-predictRisk(outgam,newdata=vetu_test,
               times=some_times,cause =1,absolute_risk_integration="exact")


```

Net of some numerical difference. 

```{r eval=FALSE, include=FALSE}
a-b
```

Why the models are scored differently?

```{r eval=FALSE, include=FALSE}
x=Score(list("GAM"=outgam,"CPH"=cph_age), #
        formula=Hist(time,status)~ 1,data=vetu_test, null.model=FALSE,
        predictRisk.args = list(coxph=list(product.limit=FALSE),
                           poisson_superlearner=list(absolute_risk_integration="approx")),
        times = 5, cause=1)

x$Brier$score

```

```{r eval=FALSE, include=FALSE}
x1=Score(list("GAM"=outgam,"CPH"=cph_age), #
        formula=Hist(time,status)~ 1,data=vetu_test,
        predictRisk.args = list(coxph=list(product.limit=TRUE),
                           poisson_superlearner=list(absolute_risk_integration="approx")),
        times = some_times, cause=1)

x2=Score(list("GAM"=outgam,"CPH"=cph_age), #
        formula=Hist(time,status)~ 1,data=vetu_test,
        predictRisk.args = list("coxph"=list(product.limit=FALSE)),
        times = some_times, cause=1)


```


## Our simulation data

```{r eval=FALSE, include=FALSE}

{
  set.seed(42)
  sample_size=300
  d <- synthesize_TD1_riskengine_data(sample_size)}

```


```{r eval=FALSE, include=FALSE}
# create id
d[,id:=1:nrow(d)]

#make factors factors
d[,c("sex",
        "value_Smoking",
        "value_Motion"):=lapply(.SD,as.factor),.SDcols=c("sex",
        "value_Smoking",
        "value_Motion")]

```



```{r eval=FALSE, include=FALSE}
lglm <- Learner_gam(covariates = c("age"),
                    intercept=FALSE)
```

```{r eval=FALSE, include=FALSE}
 out <- Superlearner(d,
                        id="id",
                        status="status_cvd",
                        stratified_k_fold=FALSE,
                        event_time = "time_cvd",
                        learners=list(lglm),
                        meta_learner_algorithms = c("glm","glmnet"),
                        nfold = 20
    )
```

```{r eval=FALSE, include=FALSE}
cox.fit  <- CSC(Hist(time_cvd,status_cvd)~age,data=d)

```

We estimate the correct coefficient.

```{r eval=FALSE, include=FALSE}
coef(out$superlearner[[1]]$learners_fit)[['age']]
#0.05508513

coef(cox.fit)
#0.05508513

```

Strange out-of sample evaluation. Sligthly different predicted risks.

```{r eval=FALSE, include=FALSE}

d_test <- synthesize_TD1_riskengine_data(floor(sample_size*30/70))

# create id
# d_test[,id:=1:nrow(d_test)]

#make factors factors
d_test[,c("sex",
        "value_Smoking",
        "value_Motion"):=lapply(.SD,as.factor),.SDcols=c("sex",
        "value_Smoking",
        "value_Motion")]

```

```{r eval=FALSE, include=FALSE}
tmp=sort(d$time_cvd)

sf.csc = predict(cox.fit,newdata=d_test[1],times=tmp,cause=1,type="survival",product.limit = FALSE)
sf.pwch = predict(out,newdata=d_test[1],times=tmp,cause=1)

sum(sf.csc$survival-sf.pwch$survival_function)
#3.004581e-06

```


```{r eval=FALSE, include=FALSE}

pr.csc = predict(cox.fit,newdata=d_test[1],times=tmp,cause=1,product.limit = FALSE)
pr.pwch.exact = predict(out,newdata=d_test[1],times=tmp,cause=1, absolute_risk_integration_method = "exact")
pr.pwch.approx = predict(out,newdata=d_test[1],times=tmp,cause=1, absolute_risk_integration_method = "approx")

# very different
data.table(sum_of_differences_exact=sum(pr.pwch.exact$absolute_risk-pr.csc$absRisk),

# numerically the same
sum_of_differences_approx=sum(pr.pwch.approx$absolute_risk-pr.csc$absRisk))

```

```{r eval=FALSE, include=FALSE}

max(a$absRisk-b$absRisk)
sum(a$absRisk-z$absolute_risk)
sum(w$absolute_risk-z$absolute_risk)
```


```{r eval=FALSE, include=FALSE}
# This is different for unobserved times
predictRisk(out,newdata=d_test[1:2,],times = c(5,10),cause=1)
predictRisk(cox.fit,newdata=d_test[1:2,],times = c(5,10),cause=1)

# This is different for observed times
tmp=sort(d$time_cvd)
predictRisk(out,newdata=d_test[1:2,],times = tmp[4],cause=1)
predictRisk(cox.fit,newdata=d_test[1:2,],times = tmp[4],cause=1,product.limit=FALSE)

comparables <- list("glm"=out,"cox"=cox.fit)

# the two cox score similar and the pwch scores differently. 

b_score_1 = Score(comparables, #
    formula=Hist(time_cvd,status_cvd)~ 1,
      predictRisk.args=list("CauseSpecificCox"=list(product.limit=FALSE),
                            "poisson_superlearner"=list(absolute_risk_integration="approx")
                            ),
    data=d_test,times=tmp, cause=1)

```

### Issues with glmnet for small samples

The algorithm struggles in converging for few data. Here an example. We need to give it a smaller treshold.


Example with no threshold .

```{r eval=FALSE, include=FALSE}
lglmnet <- Learner_glmnet(covariates = c("age"),
                    intercept=FALSE,
                    cross_validation=FALSE,
                    lambda=0)
```

```{r eval=FALSE, include=FALSE}
 out <- Superlearner(d,
                        id="id",
                        status="status_cvd",
                        stratified_k_fold=FALSE,
                        event_time = "time_cvd",
                        learners=list(lglmnet),
                        meta_learner_algorithms = c("glm","glmnet"),
                        nfold = 20
    )
```

```{r eval=FALSE, include=FALSE}
coef(out$superlearner[[1]]$learners_fit)['age',]
#0.05231385
```

Example with threshold.

```{r eval=FALSE, include=FALSE}
lglmnet <- Learner_glmnet(covariates = c("age"),
                    intercept=FALSE,
                    cross_validation=FALSE,
                    thresh =1e-14,
                    lambda=0)
```

```{r eval=FALSE, include=FALSE}
 out <- Superlearner(d,
                        id="id",
                        status="status_cvd",
                        stratified_k_fold=FALSE,
                        event_time = "time_cvd",
                        learners=list(lglmnet),
                        meta_learner_algorithms = c("glm","glmnet"),
                        nfold = 20
    )
```

```{r eval=FALSE, include=FALSE}
coef(out$superlearner[[1]]$learners_fit)['age',]
#0.05508457
```


# Model application on one simulated data set 

```{r eval=FALSE, include=FALSE}

{
  set.seed(42)
  sample_size=5000
  d <- synthesize_TD1_riskengine_data(sample_size)}

```

Simple data pre-processing. 

```{r eval=FALSE, include=FALSE}
# create id
d[,id:=1:nrow(d)]

#make factors factors
d[,c("sex",
        "value_Smoking",
        "value_Motion"):=lapply(.SD,as.factor),.SDcols=c("sex",
        "value_Smoking",
        "value_Motion")]

```

We define the models that we are using. 

```{r eval=FALSE, include=FALSE}
lgam <- Learner_gam(covariates = c("sex", "age", "s(diabetes_duration)",
                                       "value_SBP", "s(value_LDL)", "value_HBA1C",
                                       "value_Smoking", "value_Motion", "value_Albuminuria","s(eGFR)"),
                        intercept=FALSE)

    lglm <- Learner_glmnet(covariates = c("sex", "age", "diabetes_duration",
              "value_SBP", "value_LDL", "value_HBA1C",
              "value_Smoking", "value_Motion", "value_Albuminuria","eGFR"),
                         cross_validation=FALSE,
                         lambda=0,
                         intercept=FALSE,
                         penalise_nodes=TRUE)
    
    lglm <- Learner_glmnet(covariates = c("age"),
                         cross_validation=FALSE,
                         lambda=0,
                         intercept=FALSE,
                         penalise_nodes=TRUE)

    lhal <- Learner_hal(covariates = c("sex", "age", "diabetes_duration",
              "value_SBP", "value_LDL", "value_HBA1C",
              "value_Smoking", "value_Motion", "value_Albuminuria","eGFR"),
              intercept=FALSE,
                     max_degree=2,
                     smoothness_orders = 1,
                     num_knots=c(100, 75),
                     penalise_nodes=FALSE)

lbenchmark <- Learner_glmnet(covariates = c("sex", "age", "diabetes_duration",
              "value_SBP", "value_LDL", "value_HBA1C",
              "value_Smoking", "value_Motion", "value_Albuminuria","eGFR_log2:age_cut"),
                         cross_validation=FALSE,
                         lambda=0,
                         intercept=FALSE,
                         penalise_nodes=TRUE)
      

    null_model <- Learner_glmnet(
      cross_validation=FALSE,
      lambda=0,
      intercept=FALSE,
      penalise_nodes=TRUE)

```

```{r eval=FALSE, include=FALSE}

learners_list = list(lglm, lgam,lhal)

 out <- Superlearner(d,
                        id="id",
                        status="status_cvd",
                        stratified_k_fold=FALSE,
                        event_time = "time_cvd",
                        learners=learners_list,
                        nodes = seq(5,35,5),
                        min_depth=length(learners_list),
                        meta_learner_algorithms = c("glm"),
                        nfold = 20
    )

```


```{r}
lglm <- Learner_glmnet(covariates = c("age"),
                   cross_validation=FALSE,
                   lambda=0,
                   intercept=FALSE,
                   penalise_nodes=TRUE)

lhal <- Learner_hal(covariates = c("sex", "age", "diabetes_duration",
        "value_SBP", "value_LDL", "value_HBA1C",
        "value_Smoking", "value_Motion", "value_Albuminuria","eGFR"),
        intercept=FALSE,
               max_degree=2L,
               num_knots=c(10, 5),
               penalise_nodes=FALSE)
    
    
    
learners_list = list(lglm,lhal)

out <- Superlearner(d,
                  id="id",
                  status="status_cvd",
                  stratified_k_fold=FALSE,
                  event_time = "time_cvd",
                  learners=learners_list,
                  nodes = seq(5,35,5),
                  min_depth=length(learners_list),
                  meta_learner_algorithms = c("glm"),
                  nfold = 5
)
    
```



```{r eval=FALSE, include=FALSE}

# nodes <- unique(sort(sample(d$time_cvd,600)))

# nodes <- unique(sort(c(nodes, d$time_cvd[d$time_cvd>max(nodes)])))

outgam <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lgam),
                    number_of_nodes = 15,
                    # nodes =  nodes,
                    meta_learner_algorithms = c("glm_ml_1"),
                    # variable_transformation=variable_transformations,
                    nfold = 3
)
```


```{r eval=FALSE, include=FALSE}
outglm <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lglm),
                    number_of_nodes =  15,
                    meta_learner_algorithms = c("glm_ml_2"),
                    
                    # variable_transformation=variable_transformations,
                    nfold = 5
)
```

```{r eval=FALSE, include=FALSE}
coef(out$superlearner[[1]]$meta_learner_fit)
```

```{r eval=FALSE, include=FALSE}
out$meta_learner_cross_validation
```

```{r eval=FALSE, include=FALSE}

d_test <- synthesize_TD1_riskengine_data(floor(sample_size*30/70))

# create id
# d_test[,id:=1:nrow(d_test)]

#make factors factors
d_test[,c("sex",
        "value_Smoking",
        "value_Motion"):=lapply(.SD,as.factor),.SDcols=c("sex",
        "value_Smoking",
        "value_Motion")]
```

```{r eval=FALSE, include=FALSE}
    # cox.fit  <- CSC(Hist(time_cvd,status_cvd)~value_Motion+value_Smoking+eGFR+value_Albuminuria+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex,data=d,cause=1)


    cox.fit  <- CSC(Hist(time_cvd,status_cvd)~1,data=d,cause=1)

    comparables <- list("sl"=out,"glm"=outglm,"gam"=outgam,"cox"=cox.fit)
    
    b_score_1 = Score(comparables, #
    formula=Hist(time_cvd,status_cvd)~ 1,data=d_test,times=c(5,10), cause=1)
    
    b_score_2 = Score(comparables, #
    formula=Hist(time_cvd,status_cvd)~ 1,data=d_test,times=c(5,10), cause=2)

```


