---
title: "Manuscript Replication Material"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Manuscript Replication Material}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tmlensemble)
library(survival)
library(prodlim)
```

```{r}
# load the simulator 

synthesize_TD1_riskengine_data <- function(n,keep = NULL){
    #
    # empty model
    #
    sim_model <- lava::lvm()
    #
    # covariates
    #
    lava::distribution(sim_model,~sex) <- lava::binomial.lvm(p=0.462559634930512)
    lava::distribution(sim_model,~age) <- lava::normal.lvm(mean=42.2853710165578,sd=16.1358967048911)
    # this parameter seems to affect the age range:
    lava::covariance(sim_model,~age) <- 40
    lava::distribution(sim_model,~diabetes_duration) <- lava::normal.lvm(mean=-0.721917658912036,sd=11.8217218392295)
    lava::distribution(sim_model,~value_SBP) <- lava::normal.lvm(mean=113.015053986869,sd=15.1994730394504)
    lava::distribution(sim_model,~value_LDL) <- lava::normal.lvm(mean=2.34380835266364,sd=0.78907040283751)
    lava::distribution(sim_model,~value_HBA1C) <- lava::normal.lvm(mean=66.1642378865059,sd=15.1450866080122)
    lava::distribution(sim_model,~log2_eGFR_young) <- lava::normal.lvm(mean=-10.7725954841897,sd=1.85777815372659)
    lava::distribution(sim_model,~log2_eGFR_old) <- lava::normal.lvm(mean=-7.16808301118136,sd=0.295067430820889)
    lava::distribution(sim_model,~value_Smoking) <- lava::binomial.lvm(p=0.00406502756222992)
    lava::distribution(sim_model,~value_Motion) <- lava::binomial.lvm(p=0.647591503001192)
    # albuminuria
    sim_model <- lava::categorical(sim_model,~value_Albuminuria,labels=c('Normal','Micro','Macro'),K=3,p=c(0.839037544077992,0.121344119477287))
    lava::transform(sim_model, value_AlbuminuriaMicro~value_Albuminuria) <- function(x){1*(c(x)=='Micro')}
    lava::transform(sim_model, value_AlbuminuriaMacro~value_Albuminuria) <- function(x){1*(c(x)=='Macro')}
    #
    # outcome
    #
    # latent event times: .1 is cvd .2 death without cvd .0 is censoring
    lava::distribution(sim_model,~time.event.1) <- lava::coxWeibull.lvm(scale=0.00344127508109558,shape=1.1383691898171)
    lava::distribution(sim_model,~time.event.2) <- lava::coxWeibull.lvm(scale=8.8338885667654e-05,shape=1.49710154034773)
    lava::distribution(sim_model,~time.event.0) <- lava::coxWeibull.lvm(scale=0.0105086718102263,shape=2.2578792372401)
    # observed event time
    sim_model <- lava::eventTime(sim_model,time_cvd ~ min(time.event.0=0, time.event.1=1, time.event.2=2),'status_cvd')
    # uncensored event time
    sim_model <- lava::eventTime(sim_model,uncensored_time_cvd ~ min(time.event.1=1, time.event.2=2),'uncensored_status_cvd')
    #
    # structural equations
    #
    ## lava::regression(sim_model,age~sex) <- c(-0.605030442652732)
    lava::regression(sim_model,age~sex) <- c(-0.105030442652732)
    lava::regression(sim_model,diabetes_duration~age+sex) <- c(0.461453096059993,1.8834280372868)
    lava::regression(sim_model,value_SBP~diabetes_duration+age+sex) <- c(0.173711840258398,0.408090766933998,-5.26413576607228)
    lava::regression(sim_model,value_LDL~value_SBP+diabetes_duration+age+sex) <- c(0.0017997232707461,-0.00215485947423231,-0.00194439878409113,-0.0247941295129788)
    lava::regression(sim_model,value_HBA1C~value_LDL+value_SBP+diabetes_duration+age+sex) <- c(2.41425803932559,-0.0366520181671025,0.105793793763282,-0.113517351234731,1.14432850532257)
    lava::regression(sim_model,log2_eGFR_young~value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.0378605757282762,0.369860071363398,-0.000381556996934381,0.0759866855173213,-0.000921667597434942,0.0109443806788427,0.173985375155216,-0.05148192662188)
    # Fixme: it is odd to let the log2_eGFR_old depend on log2_eGFR_young
    lava::regression(sim_model,log2_eGFR_old~log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.998329448050345,0.13212871605835,0.627054487201391,-0.0017321066682523,-0.0168673394900242,0.000656653716889479,0.00231859330125351,0.0105241259989393,0.0666215469676934)
    lava::regression(sim_model,value_Smoking~log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.573224437209381,-0.533358383827049,0.417778953216348,0.811962094950751,0.0258095967578417,0.128112260674147,-0.0102291702067243,-0.00575108475194562,0.00529875681026287,-0.130756787272673)
    lava::regression(sim_model,value_Motion~value_Smoking+log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.31142087785739,-0.134144429182559,-0.0862952685994639,-0.121332306983678,-0.334529806176446,-0.0100110418345244,-0.138145721076664,0.00690508520692886,-0.0080292376964733,-0.00517733988644732,0.070748099701443)
    lava::regression(sim_model,time.event.0~value_Motion+value_Smoking+log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.0389267779620385,-0.106325304898881,-0.137101392329058,-0.118521604228187,-0.0330093143101979,0.0763711547050245,-0.00617576911315076,0.0509424048213361,-0.00757123095797519,-0.0125850223581346,-0.0130146695831196,-0.029633759327063)
    lava::regression(sim_model,time.event.1~value_Motion+value_Smoking+log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.176043482185139,0.311354666798778,0.442184803104845,0.480004039250349,0.458766977417934,0.760940706393913,0.0143694959616522,0.0701435790868005,0.00607734573133048,0.00980515944727106,0.0412496430494417,-0.224377188536924)
    lava::regression(sim_model,time.event.2~value_Motion+value_Smoking+log2_eGFR_old+log2_eGFR_young+value_AlbuminuriaMicro+value_AlbuminuriaMacro+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex) <- c(-0.410553409305998,0.833815403131808,0.3289175620287,0.313118375461698,0.225523885897843,0.604004891420425,0.0102472549995989,0.0287051547482761,0.00156423178426314,0.00678536173681053,0.0803874049498844,-0.276186036048047)
    # construction of eGFR

    # browser()
    lava::transform(sim_model,eGFR~log2_eGFR_old+log2_eGFR_young+age) <- function(x){as.numeric(x[["age"]]<40)*100*2^{x[["log2_eGFR_young"]]}+as.numeric(x[["age"]]>40)*100*2^{x[["log2_eGFR_old"]]}}
    # simulate from model
    d <- lava::sim(sim_model,n)
    data.table::setDT(d)
    d[,log2_eGFR_old := NULL]
    d[,log2_eGFR_young := NULL]
    d[,value_AlbuminuriaMicro := NULL]
    d[,value_AlbuminuriaMacro := NULL]
    d[,time.event.0 := NULL]
    d[,time.event.1 := NULL]
    d[,time.event.2 := NULL]
    if (length(keep)>0) d[,keep,with=FALSE]
    else d[]
}

```

# Model application on one simulated data set 

```{r}

{
  set.seed(42)
  sample_size=1000
  d <- synthesize_TD1_riskengine_data(sample_size)}

```


Simple data pre-processing. 

```{r}
# create id
d[,id:=1:nrow(d)]

#make factors factors
d[,c("sex",
        "value_Smoking",
        "value_Motion"):=lapply(.SD,as.factor),.SDcols=c("sex",
        "value_Smoking",
        "value_Motion")]

```


We define the models that we are using. 

```{r}


lgam <- Learner_gam(
  covariates = c(
    "sex"
  ),
  intercept = FALSE
)


lglm <- Learner_glmnet(
  covariates = c(
    "age"
  ),
  cross_validation = FALSE,
  lambda = 0,
  intercept = FALSE,
  penalise_nodes = TRUE
)

lhal <- Learner_glmnet(
  covariates = c(
    "sex", "value_Smoking", "value_Motion", "value_Albuminuria"
  ),
  cross_validation = TRUE,
  # lambda = 0,
  intercept = FALSE,
  penalise_nodes = TRUE
)

```


```{r}

# nodes <- unique(sort(sample(d$time_cvd,600)))

# nodes <- unique(sort(c(nodes, d$time_cvd[d$time_cvd>max(nodes)])))

outgam <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lgam),
                    # nodes =  nodes,
                    meta_learner_algorithms = c("glm_ml_1"),
                    # variable_transformation=variable_transformations,
                    nfold = 3
)
```


```{r}
outglm <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lglm),
                    nodes =  sort(nodes),
                    meta_learner_algorithms = c("glm_ml_1"),
                    
                    # variable_transformation=variable_transformations,
                    nfold = 20
)
```

```{r}
out <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lgam,lglm),
                    meta_learner_algorithms = c("glm"),
                    # variable_transformation=variable_transformations,
                    nfold = 3
)

```
```{r}
coef(out$superlearner[[1]]$meta_learner_fit)
```

```{r}
out$meta_learner_cross_validation
```

```{r}

d_test <- synthesize_TD1_riskengine_data(floor(sample_size*30/70))

# create id
# d_test[,id:=1:nrow(d_test)]

#make factors factors
d_test[,c("sex",
        "value_Smoking",
        "value_Motion"):=lapply(.SD,as.factor),.SDcols=c("sex",
        "value_Smoking",
        "value_Motion")]
```

```{r}
    # cox.fit  <- CSC(Hist(time_cvd,status_cvd)~value_Motion+value_Smoking+eGFR+value_Albuminuria+value_HBA1C+value_LDL+value_SBP+diabetes_duration+age+sex,data=d,cause=1)


    bchm.fit  <- CSC(Hist(time_cvd,status_cvd)~sex,data=d)
    bchm.fit1  <- coxph(Surv(time_cvd,status_cvd==1)~sex,data=d,x=TRUE)
    comparables <- list(bchm.fit,bchm.fit1,"gam"=outgam)
    
    b_score = Score(comparables, #
    formula=Hist(time_cvd,status_cvd)~ 1,data=d_test,times=sample(d$time_cvd, 3), cause=1)

```


```{r}
p1 <- predict(bchm.fit,newdata=d_test[1:2,],
        times=sort(d$time_cvd)[1],cause =1,type="survival")


p2 <- predict(outgam,newdata=d_test[1:2,],
        times=sort(d$time_cvd)[1],cause =1,type="survival")

predictRisk(bchm.fit,newdata=d_test[1,],
        times=sort(d$time_cvd)[1:9],cause =1)

predictRisk(bchm.fit1,newdata=d_test[1,],
        times=sort(d$time_cvd)[1:9],cause =1) - predictRisk(outgam,newdata=d_test[1,],
        times=sort(d$time_cvd)[1:9],cause =1)

p2 <- predict(outgam,newdata=d_test[1:2,],
        times=quantile(d$time_cvd, c(.25,.5)),cause =1)

p3 <- predictRisk(nullm,newdata=d_test[1:9,],
        times=quantile(d$time_cvd, c(.25,.5)),cause =1)

p4 <- predictRisk(outgam,newdata=d_test1:9,,
        times=quantile(d$time_cvd, c(.25,.5)),cause =1)


```


```{r}
b_score$Brier$score[,.SD,.SDcols=c("model","Brier")]
```

The hyper-parameters tuning suggested that for applying the HAL, we should try to bin the covariates in a meaningful way. In practice, the covariates are normally distributed and the experiments showed that cutting them with a finer partition within a mean plus minus two times the standard deviation type of cut improves the fitting.


We gain this type of knowledge from the simulations but a practitioner would know about this.

```{r}

lbound_1 = ceiling(mean(d$age)-2*sd(d$age))
ubound_1 = ceiling(mean(d$age)+2*sd(d$age))

age_transform = paste0("age_cut ~ sl_cut(age,breaks=c(0,",lbound_1,",",ubound_1,",100))")

lbound_2 = ceiling(mean(d$value_SBP)-2*sd(d$value_SBP))
ubound_2 = ceiling(mean(d$value_SBP)+2*sd(d$value_SBP))

value_SBP_transform = paste0("value_SBP_cut ~ sl_cut(value_SBP,breaks=c(0,",lbound_2,",",ubound_2,",160))")


lbound_3 = ceiling(mean(d$value_HBA1C)-2*sd(d$value_HBA1C))
ubound_3 = ceiling(mean(d$value_HBA1C)+2*sd(d$value_HBA1C))

value_HBA1C_transform = paste0("value_HBA1C_cut ~ sl_cut(value_HBA1C,breaks=c(0 , ",lbound_3,",",ubound_3,",100))")


lbound_4 = ceiling(mean(d$value_LDL)-2*sd(d$value_LDL))
ubound_4 = ceiling(mean(d$value_LDL)+2*sd(d$value_LDL))

value_LDL_transform = paste0("value_LDL_cut ~ sl_cut(value_LDL,breaks=c(-2,",lbound_4,",",ubound_4,",10))")


lbound_5 = ceiling(mean(log(d$eGFR))-2*sd(log(d$eGFR)))
ubound_5 = ceiling(mean(log(d$eGFR))+2*sd(log(d$eGFR)))

value_egfr_transform = paste0("log_eGFR_cut ~ sl_cut(log(eGFR),breaks=c(-5,",lbound_5,",",ubound_5,",10))")


lbound_6 = ceiling(mean((d$diabetes_duration))-2*sd((d$diabetes_duration)))
ubound_6 = ceiling(mean((d$diabetes_duration))+2*sd((d$diabetes_duration)))

value_diabetes_transform = paste0("diabetes_duration_cut ~ sl_cut(diabetes_duration,breaks=c(0,",lbound_6,",",ubound_6,",80))")

variable_transformations <- c(
  age_transform,
  value_SBP_transform,
  value_HBA1C_transform,
  value_LDL_transform,
  value_egfr_transform,
  value_diabetes_transform
)

```



We now define the super-learner.

```{r}

out <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lgam,lglm,lglm_cat),
                    nodes =  c(seq(1,20,5)),
                    meta_learner_algorithms = c("glm"),
                    variable_transformation=variable_transformations,
                    nfold = 5
)

```
```{r}
coef(out$superlearner[[1]]$meta_learner_fit)
```
```{r}
out1 <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lglm),
                    nodes =  seq(1,20,5),
                    meta_learner_algorithms = c("glm_ml_1"),
                    variable_transformation=variable_transformations,
                    nfold = 20
)
```

```{r}
out2 <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lgam),
                    nodes =  seq(1,20,5),
                    meta_learner_algorithms = c("glm_ml_1"),
                    variable_transformation=variable_transformations,
                    nfold = 20
)
```

```{r}
out3 <- Superlearner(d,
                    id="id",
                    status="status_cvd",
                    stratified_k_fold=FALSE,
                    event_time = "time_cvd",
                    learners=list(lglmnet),
                    nodes =  seq(1,20,5),
                    meta_learner_algorithms = c("glm_ml_1"),
                    variable_transformation=variable_transformations,
                    nfold = 20
)
```

```{r}

d_test <- synthesize_TD1_riskengine_data(1000)

# create id
d_test[,id:=1:nrow(d_test)]

#make factors factors
d_test[,c("sex",
        "value_Smoking",
        "value_Motion"):=lapply(.SD,as.factor),.SDcols=c("sex",
        "value_Smoking",
        "value_Motion")]
```

```{r}
    cox.fit  <- CSC(Hist(time_cvd,status_cvd)~age+sex+diabetes_duration+value_SBP+value_LDL+value_HBA1C+value_Smoking+value_Motion+value_Albuminuria+eGFR,data=d,cause=1)

    b_score = Score(list("cox"=cox.fit,"sl"=out,"glm"=out1,"gam"=out2,"glmnet"=out3),
    formula=Hist(time_cvd,status_cvd)~1,data=d_test,times=quantile(d$time_cvd, c(.25,.5,.75)), cause=1)

```


```{r}
b_score
```






