---
title: "Replicate Cox with simple survival models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicate Cox with simple survival models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tmlensemble)
library(survival)
```


Let us consider a simple survival setting

$$
O=(L_1, \widetilde{T},  {\Delta}),
$$

where we have a covariate $L_1 \in \{0,1\}$, the survival time $\widetilde{T}=\min(T,C)$, where $T\in \mathbb{R}^+$ is the time to event and $C \in \mathbb{R}^+$ represents the censoring random variable. $\Delta$ indicates censoring when equal to 0 and event when equal to 1.

We simulate data assuming a proportional model for the hazard 

$$
\lambda(t | \ell) = \lambda_0 t \exp(\beta_0  \mathbf{1}(\ell=0)+ \beta_1 \mathbf{1}(\ell=1)).
$$
In our example we will assume $\lambda_0 = 0.5$, $\beta_1 = 0.1$, and $\beta_0 = 0.25$. The censoring random variable follows a uniform distribution between $0$ and $4.5$.

We will simulate $i=1, \ldots, n$ observations of $O$, namely 


$$
O_i = (L_{1,i}, \widetilde{T}_i,  {\Delta}_i),
$$


```{r}

inverse_sf<- function(lambda, beta, X1, u){
  
  out <- -log(u)/(lambda*exp(beta*X1))
  
  return(out)
}


```

```{r}

n <- 100000

{
  set.seed(1)
  
  sample0 <- 1-runif(as.integer(n/2), min=0, max= 1)

  sample1 <- 1-runif(as.integer(n/2), min=0, max= 1)
  
  
  }

```

```{r}


times_0 <- inverse_sf(u=sample0,
                      lambda=0.5,
                      beta=0.25,
                      X1=0)

times_1 <- inverse_sf(u=sample1,
                      lambda=0.5,
                      beta=0.1,
                      X1=1)

time = c(times_0,times_1)
max(time)

{
  set.seed(1)
  cens <- runif(n=length(time),
              min=0,
              max=4.5)
  
  }

tmp <- time
tmp[time > cens] <- cens[time > cens]

dt <- data.table(
  id=1:length(dt),
  time = tmp,
  covariate = as.factor(c(rep(0,length(times_0)),
                rep(1, length(times_1)))),
  event = 1-as.numeric(time > cens)
  
  
)

table(dt$event)
```



```{r}

cox_model <- coxph(Surv(time, event) ~ covariate, dt)

summary(cox_model)
```


