---
title: "Replicate Cox with simple survival models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicate Cox with simple survival models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tmlensemble)
library(survival)
library(data.table)
```


Let us consider a simple survival setting

$$
O=(L_1, \widetilde{T},  {\Delta}),
$$

where we have a covariate $L_1 \in \{0,1\}$, the survival time $\widetilde{T}=\min(T,C)$, where $T\in \mathbb{R}^+$ is the time to event and $C \in \mathbb{R}^+$ represents the censoring random variable. $\Delta$ indicates censoring when equal to 0 and event when equal to 1.

We simulate data assuming a proportional model for the hazard 

$$
\lambda(t | \ell) = \lambda_0 t \exp(\beta_0  \mathbf{1}(\ell=0)+ \beta_1 \mathbf{1}(\ell=1)).
$$
In our example we will assume $\lambda_0 = 0.5$, $\beta_1 = 0.1$, and $\beta_0 = 0.25$. The censoring random variable follows a uniform distribution between $0$ and $4.5$.

We will simulate $i=1, \ldots, n$ observations of $O$, namely 


$$
O_i = (L_{1,i}, \widetilde{T}_i,  {\Delta}_i),
$$


```{r}

sf <- function(t,lambda, beta, X1){
  
  out <- exp(-lambda*exp(beta*X1)*t)
  return(out)
}


inverse_sf<- function(lambda, beta, X1, u){
  
  out <- -log(u)/(lambda*exp(beta*X1))
  
  return(out)
}


```

```{r}

n <- 300

{
  set.seed(1)
  
  sample <- 1-runif(as.integer(n), min=0, max= 1)
  # sample0 <- 1-runif(as.integer(n/2), min=0, max= 1)
  # sample1 <- 1-runif(as.integer(n/2), min=0, max= 1)
  X <- rnorm(n)
  
  }

```

```{r}

# 
# times_0 <- inverse_sf(u=sample0,
#                       lambda=0.5,
#                       beta=0.25,
#                       X1=0)
# 
# times_1 <- inverse_sf(u=sample1,
#                       lambda=0.5,
#                       beta=0.1,
#                       X1=1)
# 
# time = c(times_0,times_1)

time<- inverse_sf(u=sample,
                      lambda=0.5,
                      beta=0.1,
                      X1=X)

max(time)

{
  set.seed(1)
  cens <- runif(n=length(time),
              min=0,
              max=4.5)
  
  }

tmp <- time
tmp[time > cens] <- cens[time > cens]


dt <- data.table(
  id=1:length(tmp),
  time = tmp,
  covariate = X,
  status = 1-as.numeric(time > cens)
  
  
)

table(dt$status)
```



```{r}

cox_model <- coxph(Surv(time, status) ~ covariate, dt)

summary(cox_model)
```




```{r}
data_pp <- DataPreProcessing(data=dt,
                              id="id",
                              status="status",
                              setting = "survival",
                              covariates,
                              treatment,
                              # nodes= seq(1,5,by=.4),
                              event="time")
  
  
```







```{r}
model_pois <- glm("deltaij ~ covariate-1+factor(node)+offset(log(tij))",
                  data = data_pp$data,
                  family = poisson(link = "log"))


summary(model_pois)

```


```{r}

tmp <- data.frame(node=sort(unique(data_pp$data$node)),
                  tij=1,
                  covariate=1
                  )


pw_constant_hazard <- predict(model_pois,
                              newdata=tmp,
                              type = "response")


tmp[,"sf_hat"] = exp(-cumsum(pw_constant_hazard))
  

tmp<- tmp[pw_constant_hazard<=1,]
  
```

```{r}
pw_constant_hazard[2]
```



```{r}


sf(tmp$node[1],
   lambda=0.5,
   beta=0.1,
   X1=1)



plot(tmp$node,tmp$sf_hat)
```





