---
title: "Piecewise constant hazards for modelling survival"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicate Cox with simple survival models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tmlensemble)
library(survival)
library(data.table)
```


# Simulation set-up

Let us consider a simple survival setting

$$
O=(L_1, \widetilde{T},  {\Delta}),
$$

where we have a numeric covariate $L_1$, the survival time $\widetilde{T}=\min(T,C)$, where $T\in \mathbb{R}^+$ is the time to event and $C \in \mathbb{R}^+$ represents the censoring random variable. $\Delta$ indicates censoring when equal to 0 and event when equal to 1.

We simulate data assuming a proportional model for the hazard 

$$
\lambda(t | \ell) = \lambda_0 t \exp(\beta L_1).
$$
In our example we will assume $\lambda_0 = 0.5$, $\beta = 0.1$. The censoring random variable follows a uniform distribution between $0$ and $4.5$.

We will simulate $i=1, \ldots, n$ observations of $O$, namely 


$$
O_i = (L_{1,i}, \widetilde{T}_i,  {\Delta}_i),
$$


```{r}

sf <- function(t,lambda, beta, X1){
  
  out <- exp(-lambda*exp(beta*X1)*t)
  return(out)
}


inverse_sf<- function(lambda, beta, X1, u){
  
  out <- -log(u)/(lambda*exp(beta*X1))
  
  return(out)
}


```

```{r}

n <- 200

{
  set.seed(1)
  
  sample <- 1-runif(as.integer(n), min=0, max= 1)
  X <- rnorm(n)
  
  }

```

```{r}



time<- inverse_sf(u=sample,
                      lambda=0.5,
                      beta=0.1,
                      X1=X)

max(time)

{
  set.seed(1)
  cens <- runif(n=length(time),
              min=0,
              max=5.5)
  
  }

tmp <- time
tmp[time > cens] <- cens[time > cens]


dt <- data.table(
  id=1:length(tmp),
  time = tmp,
  covariate = X,
  status = 1-as.numeric(time > cens)
  
  
)

table(dt$status)
```
# Data application

## Replicating the Cox model


```{r}

cox_model <- coxph(Surv(time, status) ~ covariate, dt)

summary(cox_model)
```




```{r}
data_pp <- DataPreProcessing(data=dt,
                              id="id",
                              status="status",
                              setting = "survival",
                              covariates,
                              treatment,
                              event="time")
  
  
```


```{r}
model_pois <- glm("deltaij ~ covariate-1+factor(node)+offset(log(tij))",
                  data = data_pp$data,
                  family = poisson(link = "log"))


summary(model_pois)

```

## Different choices of the grid


In this Section we check out that we model the survival function correctly when I change the grid span. In particular, we compare the grid constructed on the observed data points, a grid with step one and a grid with step $0.5$.

### Sequence with unit distant points

```{r}
data_pp_1 <- DataPreProcessing(data=dt,
                              id="id",
                              status="status",
                              setting = "survival",
                              covariates,
                              treatment,
                              nodes = seq(0,6,1),
                              event="time")


model_pois_1 <- glm("deltaij ~ covariate-1+factor(node)+offset(log(tij))",
                  data = data_pp_1$data,
                  family = poisson(link = "log"))


```

### Sequence with 0.5 distant points

```{r}
data_pp_05 <- DataPreProcessing(data=dt,
                              id="id",
                              status="status",
                              setting = "survival",
                              covariates,
                              treatment,
                              nodes = seq(0,6,0.5),
                              event="time")


model_pois_05 <- glm("deltaij ~ covariate-1+factor(node)+offset(log(tij))",
                  data = data_pp_05$data,
                  family = poisson(link = "log"))


```

## Check out the survival function behavior

```{r}

tmp <- data.frame(node=factor(sort(unique(data_pp$data$node))),
                  tij=1,
                  covariate=0.5
                  )


pw_constant_hazard <- predict(model_pois,
                              newdata=tmp,
                              type = "response")


tmp[,"sf_hat"] = exp(-cumsum(pw_constant_hazard*diff(c(sort(unique(data_pp$data$node)),1))))



tmp_1 <- data.frame(node=factor(sort(unique(data_pp_1$data$node))),
                  tij=1,
                  covariate=0.5
                  )

pw_constant_hazard_1 <- predict(model_pois_1,
                              newdata=tmp_1,
                              type = "response")


tmp_1[,"sf_hat"] = exp(-cumsum(pw_constant_hazard_1*1))
  

tmp_05 <- data.frame(node=factor(sort(unique(data_pp_05$data$node))),
                  tij=1,
                  covariate=0.5
                  )

pw_constant_hazard_05 <- predict(model_pois_05,
                              newdata=tmp_05,
                              type = "response")


tmp_05[,"sf_hat"] = exp(-cumsum(pw_constant_hazard_05*0.5))


```





```{r}
xtrue <- seq(0,6,0.001)
ytrue <- sf(xtrue,
            lambda=.5,
            beta = 0.1,
            X1 = 0.5)

plot(xtrue,
     ytrue,
     type="line",
     xlab="Time", ylab="Survival Function")
lines(sort(unique(data_pp$data$node)),c(tmp$sf_hat), col="red",type="S")
lines(seq(0,6,by=1),c(1,tmp_1$sf_hat), col="green",type="S")
lines(seq(0,5.5,by=.5),c(1,tmp_05$sf_hat), col="blue",type="S")
legend(x = "topright", 
       title="Fitted Curve",  
       legend=c("Observed grid", "Step 1", "Step 0.5"),  
       fill = c("red","green","blue")) 




```

## Different learners 

In this Section we repeat the same experiment but we fit different learners to the data.













