---
title: "Piecewise constant hazards for modelling survival"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicate Cox with simple survival models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tmlensemble)
library(survival)
library(data.table)
library(reticulate)
library(keras)
library(xgboost)
library(glmnet)
library(rpart)
```


```{r}
seed=1
use_condaenv("C:\\Users\\pwt887\\AppData\\Local\\anaconda3\\envs\\pytmle")
Sys.setenv(PYTHONHASHSEED = seed)
set.seed(seed)
reticulate::py_set_seed(seed)
```



# Simulation set-up

Let us consider a simple survival setting

$$
O=(L_1, \widetilde{T},  {\Delta}),
$$

where we have a numeric covariate $L_1$, the survival time $\widetilde{T}=\min(T,C)$, where $T\in \mathbb{R}^+$ is the time to event and $C \in \mathbb{R}^+$ represents the censoring random variable. $\Delta$ indicates censoring when equal to 0 and event when equal to 1.

We simulate data assuming a proportional model for the hazard 

$$
\lambda(t | \ell) = \lambda_0 t \exp(\beta L_1).
$$
In our example we will assume $\lambda_0 = 0.5$, $\beta = 0.1$. The censoring random variable follows a uniform distribution between $0$ and $4.5$.

We will simulate $i=1, \ldots, n$ observations of $O$, namely 


$$
O_i = (L_{1,i}, \widetilde{T}_i,  {\Delta}_i),
$$


```{r}

sf <- function(t,lambda, beta, X1){
  
  out <- exp(-lambda*exp(beta*X1)*t)
  return(out)
}


inverse_sf<- function(lambda, beta, X1, u){
  
  out <- -log(u)/(lambda*exp(beta*X1))
  
  return(out)
}


```

```{r}

n <- 20000

{
  set.seed(seed)
  
  sample <- 1-runif(as.integer(n), min=0, max= 1)
  X <- rnorm(n)
  
  }

```

```{r}



time<- inverse_sf(u=sample,
                      lambda=0.5,
                      beta=0.1,
                      X1=X)

max(time)

{
  set.seed(1)
  cens <- runif(n=length(time),
              min=0,
              max=5.5)
  
  }

tmp <- time
tmp[time > cens] <- cens[time > cens]


dt <- data.table(
  id=1:length(tmp),
  time = tmp,
  covariate = X,
  status = 1-as.numeric(time > cens)
  
  
)

table(dt$status)
```
# Data application

## Replicating the Cox model


```{r}

cox_model <- coxph(Surv(time, status) ~ covariate, dt)

summary(cox_model)
```




```{r eval=FALSE, include=FALSE}
if(n<1000){data_pp <- DataPreProcessing(data=dt,
                              id="id",
                              status="status",
                              setting = "survival",
                              covariates,
                              treatment,
                              event="time")
  
  

model_pois <- glm("deltaij ~ covariate-1+factor(node)+offset(log(tij))",
                  data = data_pp$data,
                  family = poisson(link = "log"))


summary(model_pois)}

```

## Different choices of the grid


In this Section we check out that we model the survival function correctly when I change the grid span. In particular, we compare the grid constructed on the observed data points, a grid with step one and a grid with step $0.5$.

### Sequence with unit distant points

```{r}
data_pp_1 <- data_pre_processing(data=dt,
                              id="id",
                              status="status",
                              # setting = "survival",
                              covariates="covariate",
                              treatment=NULL,
                              nodes = seq(0,6,1),
                              event_time="time")


model_pois_1 <- glm("deltaij ~ covariate-1+factor(node)+offset(log(tij))",
                  data = data_pp_1$data,
                  family = poisson(link = "log"))


```

### Sequence with 0.5 distant points

```{r}
data_pp_05 <- DataPreProcessing(data=dt,
                              id="id",
                              status="status",
                              setting = "survival",
                              covariates,
                              treatment,
                              nodes = seq(0,6,0.5),
                              event="time")


model_pois_05 <- glm("deltaij ~ covariate-1+factor(node)+offset(log(tij))",
                  data = data_pp_05$data,
                  family = poisson(link = "log"))


```

## Check out the survival function behavior

```{r}


if (n < 1000) {
  tmp <- data.frame(node = factor(sort(unique(
    data_pp$data$node
  ))),
  tij = 1,
  covariate = 0.5)
  
  
  pw_constant_hazard <- predict(model_pois, newdata = tmp, type = "response")
  tmp[, "sf_hat"] = exp(-cumsum(pw_constant_hazard * diff(c(sort(
    unique(data_pp$data$node)
  ), 1))))
}



tmp_1 <- data.frame(node = factor(sort(unique(
  data_pp_1$data$node
))),
tij = 1,
covariate = 0.5)

pw_constant_hazard_1 <- predict(model_pois_1, newdata = tmp_1, type = "response")


tmp_1[, "sf_hat"] = exp(-cumsum(pw_constant_hazard_1 * 1))


tmp_05 <- data.frame(node = factor(sort(unique(
  data_pp_05$data$node
))),
tij = 1,
covariate = 0.5)

pw_constant_hazard_05 <- predict(model_pois_05, newdata = tmp_05, type = "response")


tmp_05[, "sf_hat"] = exp(-cumsum(pw_constant_hazard_05 * 0.5))


```





```{r}
xtrue <- seq(0,6,0.001)
ytrue <- sf(xtrue,
            lambda=.5,
            beta = 0.1,
            X1 = 0.5)

plot(xtrue,
     ytrue,
     type="line",
     xlab="Time", ylab="Survival Function")
if(n<1000){lines(sort(unique(data_pp$data$node)),c(tmp$sf_hat), col="red",type="S")}
lines(seq(0,6,by=1),c(1,tmp_1$sf_hat), col="green",type="S")
lines(seq(0,5.5,by=.5),c(1,tmp_05$sf_hat), col="blue",type="S")
legend(x = "topright", 
       title="Fitted Curve",  
       legend=c("Observed grid", "Step 1", "Step 0.5"),  
       fill = c("red","green","blue")) 




```

## Different learners 

In this Section we repeat the same experiment but we fit different learners to the data.


### Neural Networks 

```{r}
preprocess_minmax <- function(varData) {
  X <- as.numeric(varData)
  2 * (X - min(X)) / (max(X) - min(X)) - 1
}


x_nn <- data_pp_1$data %>%
  select(covariate,node) %>%
  mutate(covariate=preprocess_minmax(covariate))%>%
    fastDummies::dummy_cols(select_columns = "node",
               remove_selected_columns = TRUE,
               remove_first_dummy = FALSE)

y_nn <- data_pp_1$data %>%
  select(deltaij,tij) 

nfeatures <- dim(x_nn)[2]


Design  <- keras::layer_input(shape = c(nfeatures), dtype = 'float32', name = 'Design') 
LogVol  <- layer_input(shape = c(1), dtype = 'float32', name = 'LogVol')

Network <- Design %>%
  layer_dense(units = 2, activation = 'tanh', name = 'layer1',  use_bias = F) %>%
  layer_dense(units = 1, activation = 'exponential', name = 'Network',  use_bias = F)

Response <- LogVol %>%
  layer_dense(units = 1,  use_bias = F, activation = k_exp, name = 'Response', trainable = FALSE,
              weights =list( array(1, dim = c(1, 1))))


# Combine the outputs of the two networks
combined_output <- layer_multiply(list(Network, Response))

# Create a model with the combined output
model <- keras_model(inputs = list(Design, LogVol), outputs = combined_output)

```

```{r}
poisson_nll <- function(y_true, y_pred, ...) {

  out <- -sum(y_true * log(y_pred) - y_pred)

  return(out)

}


model %>% compile(
  loss = 'poisson',
  optimizer = 'adam'
)

summary(model)




```

```{r}
# set hyperparameters
epochs <- 300
batch_size <- 2000
validation_split <- 0.2  # set to >0 to see train/validation loss in plot(fit)
verbose <- 1
```


```{r}
model_fit <- model %>% fit(
    list(as.matrix.data.frame(x_nn), as.matrix(log(y_nn$tij))), as.matrix((y_nn$deltaij)),
    epochs = epochs,
    batch_size = batch_size,
    validation_split = validation_split,
    verbose = verbose
  )
```

```{r}


X_nn_test = tmp_1 %>%
  select(covariate,node) %>%
  mutate(covariate=2 * (covariate - min(data_pp_1$data$covariate)) / (max(data_pp_1$data$covariate) - min(data_pp_1$data$covariate)) - 1)%>%
    fastDummies::dummy_cols(select_columns = "node",
               remove_selected_columns = TRUE,
               remove_first_dummy = FALSE) %>%
  as.matrix()
  



pw_constant_hazard_1_nn <- as.vector(model%>% predict(list(X_nn_test, as.matrix(rep(log(1),dim(X_nn_test)[1])))))



tmp_1['sf_nn'] <- exp(-cumsum(pw_constant_hazard_1_nn*1))

```


### Xgboost

```{r}

data_pp_1_xgb <- data_pp_1$data %>% mutate(node=factor(node))
                                      
#deltaij ~ covariate-1+factor(node)+offset(log(tij))


xgtrain.mf  <- model.frame(as.formula("deltaij ~ covariate-1+node+offset(log(tij))"),data_pp_1_xgb)
xgtrain.m  <- model.matrix(attr(xgtrain.mf,"terms"),data = data_pp_1_xgb)
xgtrain  <- xgb.DMatrix(xgtrain.m,label = data_pp_1_xgb$deltaij)
setinfo(xgtrain, "base_margin", log(data_pp_1_xgb$tij))


```



```{r}
xgb = xgb.train(
  nrounds = 800
  , params = list(objective="count:poisson",
                  eta=.01,
                  max_depth =2,
                  alpha=.5,
                  lambda=.5)
  , data = xgtrain
)
```

```{r}
tmp_1_xgb <-tmp_1%>% mutate(deltaij=1)
xgtrain.mf  <- model.frame(as.formula("deltaij ~ covariate-1+node+offset(log(tij))"),tmp_1_xgb)
xgtrain.m  <- model.matrix(attr(xgtrain.mf,"terms"),data = tmp_1_xgb)
xgtest  <- xgb.DMatrix(xgtrain.m,label = tmp_1_xgb$deltaij)
setinfo(xgtest, "base_margin", log(tmp_1_xgb$tij))

pw_constant_hazard_1_xgb <- predict(xgb, newdata = xgtest)

tmp_1['sf_xgb'] <- exp(-cumsum(pw_constant_hazard_1_xgb*1))
tmp_1
```
## Glmnet


```{r}

data_pp_1_glmnet <- data_pp_1$data %>% mutate(node=factor(node))

glmnet_train.mf  <- model.frame(as.formula("deltaij ~ covariate-1+node+offset(log(tij))"),data_pp_1_glmnet)
x_glmnet  <- model.matrix(attr(xgtrain.mf,"terms"),data = data_pp_1_glmnet)


offset_glmnet <- data_pp_1_glmnet %>% select(tij) %>% mutate(tij=log(tij))%>% as.matrix.data.frame()
y_glmnet <- data_pp_1_glmnet %>% select(deltaij) %>% as.matrix.data.frame()

cvfit <- cv.glmnet(x_glmnet, y_glmnet, offset = offset_glmnet, family="poisson")

glmnet_fit <- glmnet(x_glmnet, y_glmnet, offset = offset_glmnet, family="poisson",lambda = cvfit$lambda.min)


```

```{r}
tmp_1_glmnet <- tmp_1 %>% mutate(deltaij=1,
                                 tij=1)

glmnet_test.mf  <- model.frame(as.formula("deltaij ~ covariate-1+node+offset(log(tij))"),tmp_1_glmnet)

x_glmnet  <- model.matrix(attr(glmnet_test.mf,"terms"),data = tmp_1_glmnet)
newoffset_glmnet = rep(0,dim(x_glmnet)[1])
pw_constant_hazard_1_glmnet <- predict(glmnet_fit, newx = x_glmnet, newoffset= newoffset_glmnet, type = "response")
tmp_1['sf_glmnet'] <- exp(-cumsum(pw_constant_hazard_1_glmnet*1))


```

### Regression trees

```{r}

data_pp_1_rpart <- data_pp_1$data %>% mutate(node=factor(node))

rpart_train.mf  <- model.frame(as.formula("deltaij ~ covariate-1+node+offset(log(tij))"),data_pp_1_rpart)
x_rpart  <- model.matrix(attr(xgtrain.mf,"terms"),data = data_pp_1_rpart)
offset_rpart <- data_pp_1_rpart %>% select(tij) %>% mutate(tij=(tij))%>% as.matrix.data.frame()
y_rpart <- data_pp_1_glmnet %>% select(deltaij) %>% as.matrix.data.frame()

model_rpart = rpart(cbind(offset_rpart, y_rpart) ~ covariate + node ,data=data_pp_1_rpart,method="poisson")



```


```{r}
tmp_1_rpart <- tmp_1 %>% mutate(deltaij=1,
                                 tij=1)


pw_constant_hazard_1_rpart <- predict(model_rpart, tmp_1_rpart)
tmp_1['sf_rpart'] <- exp(-cumsum(pw_constant_hazard_1_rpart*1))

```



### Check out the results

```{r}
xtrue <- seq(0,6,0.001)
ytrue <- sf(xtrue,
            lambda=.5,
            beta = 0.1,
            X1 = 0.5)

plot(xtrue,
     ytrue,
     type="line",
     xlab="Time", ylab="Survival Function")
# lines(sort(unique(data_pp$data$node)),c(tmp$sf_hat), col="red",type="S")
lines(seq(0,6,by=1),c(1,tmp_1$sf_hat), col="red",type="S")
lines(seq(0,6,by=1),c(1,tmp_1$sf_nn), col="green",type="S")
lines(seq(0,6,by=1),c(1,tmp_1$sf_xgb), col="blue",type="S")
lines(seq(0,6,by=1),c(1,tmp_1$sf_glmnet), col="orange",type="S")
lines(seq(0,6,by=1),c(1,tmp_1$sf_rpart), col="pink",type="S")
legend(x = "topright", 
       title="Fitted Curve",  
       legend=c( "glm", "nn","xgboost","glmnet","rpart"),  
       fill = c("red","green","blue","orange","pink")) 


```

# Learners

```{r}
tmp2 <- data_pre_processing(
  data = dt,
  id = "id",
  status = "status",
  # covariates = c("Sex", "D", "Phase", "Age", "Source"),
  # treatment = NULL,
  nodes = seq(0,6,0.5),
  event_time = "time"
)
```


```{r}
model_pois_1 <- glm("deltaij ~ covariate-1+factor(node)+offset(log(tij))",
                  data = tmp2,
                  family = poisson(link = "log"))
```


```{r}
tmp_05 <- data.frame(node = factor(sort(unique(
  tmp2$node
))),
tij = 1,
covariate = 0.5)

pw_constant_hazard <- predict(model_pois_1, newdata = tmp_05, type = "response")


tmp_05[, "sf_hat"] = exp(-cumsum(pw_constant_hazard * 0.5))
```

```{r}
l1 <- learner_glmnet(covariates = c("covariate"),
                     treatment = NULL,
                     cross_validation=T)
```

```{r}
fit <- l1(tmp2)
```

<!-- ```{r} -->

<!-- # glmnet_test.mf  <- model.frame(as.formula("deltaij ~ covariate-1+node+offset(log(tij))"),tmp_05) -->

<!-- pw_constant_hazard_1_glmnet <- predict(fit, newx = tmp_05 %>%select(covariate,node) %>%as.matrix(), newoffset= 0, type = "response", s = "lambda.min") -->

<!-- tmp_05['sf_glmnet'] <- exp(-cumsum(pw_constant_hazard_1_glmnet*.5)) -->


<!-- tmp_1_glmnet <- tmp_1 %>% mutate(deltaij=1, -->
<!--                                  tij=.5) -->

<!-- # glmnet_test.mf  <- model.frame(as.formula("deltaij ~ covariate-1+node+offset(log(tij))"),tmp_1_glmnet) -->

<!-- x_glmnet  <- model.matrix(attr(glmnet_test.mf,"terms"),data = tmp_1_glmnet) -->
<!-- newoffset_glmnet = rep(0,dim(x_glmnet)[1]) -->
<!-- pw_constant_hazard_1_glmnet <- predict(glmnet_fit, newx = x_glmnet, newoffset= newoffset_glmnet, type = "response") -->
<!-- tmp_1['sf_glmnet'] <- exp(-cumsum(pw_constant_hazard_1_glmnet*1)) -->


<!-- ``` -->
<!-- ```{r} -->
<!-- xtrue <- seq(0,6,0.001) -->
<!-- ytrue <- sf(xtrue, -->
<!--             lambda=.5, -->
<!--             beta = 0.1, -->
<!--             X1 = 0.5) -->


<!-- plot(xtrue, -->
<!--      ytrue, -->
<!--      type="line", -->
<!--      xlab="Time", ylab="Survival Function") -->
<!-- if(n<1000){lines(sort(unique(data_pp$data$node)),c(tmp$sf_hat), col="red",type="S")} -->
<!-- lines(seq(0,5.5,by=.5),c(1,tmp_05$sf_glmnet), col="blue",type="S") -->
<!-- legend(x = "topright",  -->
<!--        title="Fitted Curve",   -->
<!--        legend=c("Observed grid", "Step 1", "Step 0.5"),   -->
<!--        fill = c("red","green","blue"))  -->
<!-- ``` -->




